---
title: "Cyber-Security Hackathon"
author: "Gjeltema"
date: "05/21/2015"
output: html_document
---
Hackathon for Research Triangle Analysts  
http://www.meetup.com/de/Research-Triangle-Analysts/events/221861919/  
R code for a basic analysis  

# Packages
```{r, warning=FALSE, message=FALSE}
# install.packages("GGally"); install.packages("ggplot2");
# install.packages("dplyr"); install.packages("magrittr");
# install.packages("tidyr");
suppressPackageStartupMessages(library("ggplot2"))
suppressPackageStartupMessages(library("dplyr"))
suppressPackageStartupMessages(library("magrittr"))
suppressPackageStartupMessages(library("tidyr"))
```

# Load data & view summary statistics
```{r loadData}
evilprog <- read.csv("evilProgram_flat.csv", stringsAsFactors = FALSE);
evilprog.orig <- evilprog;
evilprog <- evilprog.orig;
dim(evilprog); names(evilprog);
str(evilprog); # summary(evilprog);

# choose the apropriate data type for each variable
evilprog %<>% mutate(Protocol = as.factor(Protocol),
                     Source = as.factor(Source),
                     Destination = as.factor(Destination))
```

## Why so many variables? Is there even anything in them?
```{r analyzeVariables}
evilprog %>% 
  select(starts_with("Additional")) %>% 
  mutate_each(funs(factor)) %>% 
  summary();

# looks like additional info variables 9 till the last one are mostly empty, 
# except for a few observations
evilprog %>% filter(AdditionalInfo9 != "") %>% count();
evilprog %>%
  filter(AdditionalInfo9 != "") %>%
  select(num_range("AdditionalInfo",7:30)) %>% 
  distinct() %>% count();

# drop the columns that seem redundant
evilprog %<>% select(No, Time, Source, Destination, Protocol, Length, 
                     num_range("AdditionalInfo",1:8))

summary(evilprog)
```

# Let's look at the data
```{r plotData}
GGally::ggpairs(evilprog, columns = 2:6, title = "Scatter Plot")
```

## Suspiciously much traffic any time? (e.g. lots of traffic in short time span)
```{r}
ggplot(data=evilprog, aes(x=Time)) + 
  geom_histogram(binwidth=100) + 
  ggtitle("Any suspicious events when counting traffic?")
```

## Traffic volumes (length) over time
```{r}
ggplot(evilprog, aes(x=Time, y=Length, color=Protocol)) + 
  geom_point() + 
  ggtitle("Any suspicious events in the traffic volumes over time?")
```

## Traffic by type of protocol
```{r}
ggplot(data=evilprog, aes(x=Length, fill=Protocol)) + 
  geom_bar() + 
  facet_wrap(~Protocol, scales="free") + 
  theme(legend.position="none") + 
  ggtitle("Any unusual packet lengths per protocol?")
```

## Big volume of packet lengths from certain sources or to certain destinations?
```{r}
big_sizes.source <-
evilprog %>% 
  filter(Length>60) %>% 
  select(Source, Destination, Length) %>%
  group_by(Source) %>%
  summarize(total_traffic=sum(Length)) %>%
  arrange(desc(total_traffic));

big_sizes.source %>% 
ggplot(., aes(x=reorder(Source, total_traffic), y=total_traffic)) + 
  xlab("Source IP address") + geom_point() + coord_flip() +
  ggtitle("Sources of heavy traffic?")



big_sizes.destination <-
evilprog %>% 
  filter(Length>60) %>% 
  select(Source, Destination, Length) %>%
  group_by(Destination) %>%
  summarize(total_traffic=sum(Length)) %>%
  arrange(desc(total_traffic));

big_sizes.destination %>%
ggplot(., aes(x=reorder(Destination, total_traffic), y=total_traffic)) + 
  xlab("Destination IP address") + 
  geom_point() + coord_flip() + 
  ggtitle("Destinations with heavy traffic?")



big_sizes.all <-
evilprog %>% 
  filter(Length>60) %>% 
  select(Source, Destination, Length) %>%
  group_by(Source, Destination) %>%
  summarize(total_traffic=sum(Length)) %>%
  data.frame() %>%
  arrange(desc(total_traffic));
```


